---
title: "P8105 HW 1"
author: "Vanessa Dinh (vd2407)"
date: 2022-9-22
output: github_document
---

```{r load_libraries, results = FALSE}
library(tidyverse)
```

# Problem 1

```{r problem_1}
data("penguins", package = "palmerpenguins")
view(penguins)
str(penguins)
head(penguins)
tail(penguins)

names(penguins) 
ncol(penguins)
nrow(penguins)
mean(penguins$flipper_length_mm, na.rm = TRUE)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("bill_vs_flipper.png")
```

The penguins dataset has 8 columns and 344 rows. There are 8 variables which are species, island, bill length (mm), bill depth (mm), flipper_length (mm), body mass (grams), sex, and year. Species, island, and sex are factor variables while the rest are numeric. The dataset includes 3 penguin species, 3 different islands, and two sexes across years 2007 to 2009. 

The mean flipper length within the penguins dataset is 200.9152 mm.

# Problem 2

```{r problem_2_part_1, eval = TRUE}
prob_2 <- tibble(
  random_sample = rnorm(10),
  log_vec = random_sample > 0,
  char_vec = c("a","b","d","c","b","a","c","b","d","a"),
  fact_vec = factor(char_vec, levels = c("a","b","c"))
)

view(prob_2)

mean(prob_2$random_sample)
mean(prob_2$log_vec)
mean(prob_2$char_vec)
mean(prob_2$fact_vec)

```
The mean function works for the random sample and logical vectors, but returns NA for character and factor vectors. Since the logical vector codes 0 and 1 for FALSE and TRUE, the mean function takes the average of these as if they were numeric. 

```{r problem_2_part_2, results = FALSE}
as.numeric(prob_2$log_vec)
as.numeric(prob_2$char_vec)
as.numeric(prob_2$fact_vec)
```
The as.numeric function coerces a vector of any variable type to become a numeric variable. However, there is a warning attached to the character vector that displays NA since they weren't able to attach a number/order to these values. Since a factor vector is categorical, a value can be given in order of the levels defined in the factor. This does help explain what happens when I try to take the mean, since as.numeric attempts to assign a number to each observation in the vector and displays NA when it is unable to do so.

